---
title: "Iris: ANOVA, Tukey HSD, and Compact Letter Display"
author: "Lab demo"
date: today
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

## Question

Do the three *Iris* species differ in **sepal length**, and which pairs
differ from each other?

## Data

```{r}
#| label: data
data(iris)
summary(iris$Sepal.Length)
table(iris$Species)
```

## One-way ANOVA

```{r}
#| label: anova
fit <- aov(Sepal.Length ~ Species, data = iris)
summary(fit)
```

The species factor is highly significant — at least one species mean differs
from the others. ANOVA does not tell us *which* pairs differ; for that we use
Tukey's HSD post-hoc test.

## Tukey HSD

```{r}
#| label: tukey
tuk <- TukeyHSD(fit)
tuk
```

## Compact letter display

Means that share a letter are not significantly different at $\alpha = 0.05$.

```{r}
#| label: cld
library(multcompView)

cld <- multcompLetters4(fit, tuk)
cld_letters <- cld$Species$Letters
cld_df <- data.frame(
  Species = names(cld_letters),
  cld     = unname(cld_letters)
)
cld_df
```

## Plot with letters

```{r}
#| label: fig-cld
#| fig-cap: "Sepal length by *Iris* species. Means sharing a letter are not significantly different (Tukey HSD, α = 0.05)."
#| fig-width: 6
#| fig-height: 4
library(ggplot2)

y_label_pos <- max(iris$Sepal.Length) + 0.4

ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 1) +
  geom_text(
    data = cld_df,
    aes(x = Species, y = y_label_pos, label = cld),
    inherit.aes = FALSE, size = 6, fontface = "bold"
  ) +
  labs(
    y = "Sepal length (cm)",
    x = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Session info

```{r}
#| label: session
sessionInfo()
```
